---
title: "Chapter_10_Assignment"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE,autodep = TRUE)
```

```{r}
library(rethinking)
library(ggplot2)
```


## 10E1
_If an event has probability of 0.35 what are the log-odds of this event?_

The log odds are 
```{r}
log(.35/(1-.35))
```

## 10E2
_If an event has log-odds of 3.2, what is the probability of this event?_

The probability is
```{r}
logistic(3.2) # this is p/(1-p)
```

## 10E3
_Suppose that a coefficient in a logistic regression has a value of 1.7.  What does this imply about the proportional change in odds ofthe outcome_

There is a proportional increase of
```{r}
exp(1.7)
```
in the odds of the event

## 10M1
_Bionomial data can be organized as aggregated or disaggregated without any impact on inference.  The likelihood of the data does change.  Why?_

```{r}
dbinom(5,10,.5)
dbinom(rep(c(0,1),each=5),1,.5)
```

The latter does not take into account all of the different ways that the observations could have occured.

## 10M3
_Explain why the logit link is appropriate for a binomial model_

Because it allows modeling of p, the important parameter for the binomial, as a linear function of the predictors, and limits p to between 0 and 1.

## 10H1

