---
title: "Chapter_10_Assignment"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE,autodep = TRUE)
```

```{r}
library(rethinking)
library(ggplot2)
```


## 10E1
_If an event has probability of 0.35 what are the log-odds of this event?_

The log odds are 
```{r}
log(.35/(1-.35))
```

## 10E2
_If an event has log-odds of 3.2, what is the probability of this event?_

The probability is
```{r}
logistic(3.2) # this is p/(1-p)
```

## 10E3
_Suppose that a coefficient in a logistic regression has a value of 1.7.  What does this imply about the proportional change in odds ofthe outcome_

There is a proportional increase of
```{r}
exp(1.7)
```
in the odds of the event

## 10E4
_Why do Poisson regressons sometimes require an offset?_

An offset is required of the measurement interval is different for different observations.  Perhaps you are comparing transposon insertion rates in two different species and in one the reported rate is per 100kb and the other is per 10kb.

## 10M1
_Bionomial data can be organized as aggregated or disaggregated without any impact on inference.  The likelihood of the data does change.  Why?_

```{r}
dbinom(5,10,.5)
dbinom(rep(c(0,1),each=5),1,.5)
```

The latter does not take into account all of the different ways that the observations could have occured.

## 10M2
_If a coefficint in a Poisson regression has a value of 1.7, what does this imply about the change in outcome?_

For this we need to use the inverse function
```{r}
exp(1.7)
```
each unit of change in the predictor increases the lambda (aka the mean and variance) by 5.47.

## 10M3
_Explain why the logit link is appropriate for a binomial model_

Because it allows modeling of p, the important parameter for the binomial, as a linear function of the predictors, and limits p to between 0 and 1.

## 10M4
_Explain why the log link is appropriate for Poisson GLM_
The log link ensures that the outcome is positive.

## 10M5
_What would it impoly to use a logit link for a Poisson GLM?  Why might you want to_
IT implies that the mean is between 0 and 1.  This makes sense if it is an impossibilty for the event to occur more than once per unit of measurement.  So if you are measureing the number of balls per widget and each widget can only hold a single ball and you measure every widget.

## 10H1

_Use `map` to construct a quadractic approximation of the posterior distribution for m10.4.  Compare to MCMC and explain the similarities and differences_

### First set up the data and run the MCMC model
```{r}
data("chimpanzees")
d <- chimpanzees
d2 <- d
d2$recipient <- NULL
m10.4.stan <- map2stan(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left ,
    a[actor] ~ dnorm(0,10),
    bp ~ dnorm(0,10),
    bpC ~ dnorm(0,10)
  ),
  data=d2 , chains=2 , iter=2500 , warmup=500 )
```

### Now run the quadratic approximation
```{r}
m10.4.map <- map(  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a[actor] + (bp + bpC*condition)*prosoc_left ,
    a[actor] ~ dnorm(0,10),
    bp ~ dnorm(0,10),
    bpC ~ dnorm(0,10)
  ),
  data=d2)
```

### compare
```{r}
precis(m10.4.stan,depth=2)
precis(m10.4.map,depth=2)
compare(m10.4.stan,m10.4.map) # not a good idea?
coeftab(m10.4.map,m10.4.stan)
plot(coeftab(m10.4.map,m10.4.stan))
pairs(m10.4.map)
pairs(m10.4.stan)
```

The models differ the most in the posterior for a2 and a7.  These posteriors have the most non--Guassian distribution.

## 10H2
_Use WAIC to compare m10.4 to the simpler models fit to the same data_

So I guess I shoud fit the all with stan
```{r}
m10.1.stan <- map2stan(
  alist(
    pulled_left ~ dbinom(1,p),
    logit(p) <- a,
    a ~ dnorm(0,10)
  ),
  data=d2)

m10.2.stan <- map2stan(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a + bp*prosoc_left ,
    a ~ dnorm(0,10) ,
    bp ~ dnorm(0,10)
  ),
  data=d2 )

m10.3.stan <- map2stan(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a + (bp + bpC*condition)*prosoc_left ,
    a ~ dnorm(0,10) ,
    bp ~ dnorm(0,10) ,
    bpC ~ dnorm(0,10)
  ), data=d2 )

compare(m10.1.stan,m10.2.stan,m10.3.stan,m10.4.stan)
```

Conclusion: the model with an individual intercept for each actor is strongly favored.

## 10H4

__(a)__ _Model the relationship between density and percent cover.  In what way does the model do a good and bad job?_

```{r}
library(rethinking)
library(ggplot2)
data(salamanders)
head(salamanders)
qplot(x=PCTCOVER,y=SALAMAN,geom="point",data=salamanders)
msal1 <- map(alist( SALAMAN ~ dpois(lambda),
                    log(lambda) <- a + b_c*PCTCOVER,
                    a ~ dnorm(0,2),
                    b_c ~ dnorm(0,1)),
             data=salamanders)
precis(msal1,corr = TRUE)
plot(precis(msal1))                    
```

```{r}
pred.df <- data.frame(PCTCOVER=seq(0,100,1))
lambda.link <- link(msal1,pred.df)
pred.df$lambda.med <- apply(lambda.link,2,median)
pred.df <- cbind(pred.df,t(apply(lambda.link,2,PI)))
colnames(pred.df)[c(3,4)] <- c("PI.low","PI.high")
pl <- ggplot(pred.df,aes(x=PCTCOVER,y=lambda.med))
pl <- pl + geom_line()
pl <- pl + geom_ribbon(aes_string(ymin="PI.low",ymax="PI.high"),alpha=.1)
pl <- pl + geom_point(data=salamanders,aes(y=SALAMAN)) 
pl + ylab("Salamander density")
```


```{r}
msal1.stan <- map2stan(msal1,chains=4,iter=4000)
precis(msal1.stan) #pretty poor sampling, really.  Due to correlation I assume.
pairs(msal1.stan)
```


The model does a good job at low PCTCOVER and captures the average increase but does not account for the variability at high PCTCOVER.


__(b)__ _Can you improve the model by using the FORESTAGE predictor?  Try any models that may be useful.  Explain why FORESTAGE helps or does not help_

Additive model 

```{r}
qplot(x=FORESTAGE,y=SALAMAN,geom="point",data=salamanders)
qplot(x=FORESTAGE,y=PCTCOVER,geom="point",data=salamanders)
msal.add <- map(alist( SALAMAN ~ dpois(lambda),
                    log(lambda) <- a + b_c*PCTCOVER + b_age*FORESTAGE,
                    a ~ dnorm(0,2),
                    c(b_c,b_age) ~ dnorm(0,1)),
                method="Nelder-Mead",
                start = list(a=0,b_c=0,b_age=0),
             data=salamanders)
precis(msal.add,corr = TRUE) #note : this can be really variable!
plot(precis(msal.add))  

(compare.out <- compare(msal1,msal.add)) #model with FORESTAGE is a bit worse
plot(compare.out)
```

Model that includes FORESTAGE is a bit worse than only with PCTCOVER and FORESTAGE is not predicting anything.  (note that the coefficient estimates can vary a lot with different runs; this also makes me think that we are over-paramterized here).  Bottom line, this additive model does not help.

What about FORESTAGE on its own?

```{r}
msal.age <- map(alist( SALAMAN ~ dpois(lambda),
                    log(lambda) <- a +  b_age*FORESTAGE,
                    a ~ dnorm(0,2),
                    c(b_age) ~ dnorm(0,1)),
                method="Nelder-Mead",
             data=salamanders)
precis(msal.age,corr = TRUE)
plot(precis(msal.age))  

(compare.out <- compare(msal1,msal.add,msal.age)) 
plot(compare.out)
```

Model with FORESTAGE alone is much worse.

Model with interaction

```{r eval=FALSE}
msal.int <- map(alist( SALAMAN ~ dpois(lambda),
                    log(lambda) <- a + b_c*PCTCOVER + b_age*FORESTAGE + b_c_age*PCTCOVER * FORESTAGE,
                    a ~ dnorm(0,2),
                    c(b_c,b_age,b_c_age) ~ dnorm(0,1)),
             data=salamanders,
             method="Nelder-Mead",
             start=list(a=0,b_c=0,b_age=0,b_c_age=0))
precis(msal1,corr = TRUE)
plot(precis(msal1)) 
```

I cannot get this to work.
