---
title: "Germination Analysis"
author: "Julin Maloof"
date: "10/4/2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
library(rethinking)
library(ggplot2)
```

```{r subset_data, echo=FALSE, eval=FALSE}
d <- read.csv("LironData.csv")
summary(d)
with(d,table(Germination,Time))
write.csv(d[d$Time==1,],"LironDataTime1.csv",row.names = FALSE)
```

```{r read_data}
d <- read.csv("LironDataTime1.csv")
summary(d)
head(d)
d$Germination2 <- as.numeric(d$Germination)-1 #0 = no, 1 = yes
d$pen <- as.numeric(d$Species)-1 # 1 = pen
d$Temperature_b <- d$Temperature-min(d$Temperature) # baselined temperature
d.stan <- d[,grepl("Germination2|pen|Temperature_b",colnames(d))]
head(d.stan)
```

plot the data

```{r plot}
pl <- ggplot(d,aes(x=Germination,fill=Germination))
pl + geom_bar() + facet_grid(Species ~ Temperature)
```

Make an additive model

```{r model1, results='hide'}
m1 <- map2stan(alist(Germination2 ~ dbinom(1, p),
                     logit(p) <- a + bPen*pen + bTemp*Temperature_b,
                     c(a,bPen,bTemp) ~ dnorm(0,5)),
               data=d.stan,chains = 4)
```

```{r model1_results}
plot(m1)
pairs(m1)
precis(m1)
```

The PI for both coefficients are above 0 so it seems that both temperature and species have an effect

### Interaction model

```{r model2, results='hide'}
m2 <- map2stan(alist(Germination2 ~ dbinom(1, p),
                     logit(p) <- a + bPen*pen + bTemp*Temperature_b + bPen_Temp*pen*Temperature_b,
                     c(a,bPen,bTemp,bPen_Temp) ~ dnorm(0,5)),
               data=d.stan,chains = 4)
```

```{r model2_results}
plot(m2)
pairs(m2)
precis(m2)
```


### Species only model

```{r model3, results='hide'}
m3 <- map2stan(alist(Germination2 ~ dbinom(1, p),
                     logit(p) <- a + bPen*pen,
                     c(a,bPen) ~ dnorm(0,5)),
               data=d.stan,chains = 4) 
```

```{r model3_Results}
plot(m3)
pairs(m3)
precis(m3)
```

### Temperature only model

```{r model4, results='hide'}
m4 <- map2stan(alist(Germination2 ~ dbinom(1, p),
                     logit(p) <- a + bTemp*Temperature_b,
                     c(a,bTemp) ~ dnorm(0,5)),
               data=d.stan,chains = 4)
```

```{r model4_results}
plot(m4)
pairs(m4)
precis(m4)
```

### Compare models

```{r compare}
compare(m1,m2,m3,m4)
coeftab(m1,m2,m3,m4)
```

The interactive and full additive models have a lot of support; the single factor models do not.

### Get predictions

```{r}
pred.df <- data.frame(
  pen=c(0,0,1,1),
  Temperature_b=c(0,8,0,8))
pred.df
p.link <- ensemble(m1,m2,data=pred.df,n=4000)$link #using 4000 to try to get smoother estimates
pred.df$p.mean <- apply(p.link,2,mean)
pred.df$p.PI.low <- apply(p.link,2,PI)[1,]
pred.df$p.PI.high <- apply(p.link,2,PI)[2,]

#add nice plotting columns

pred.df$temperature <- pred.df$Temperature_b+min(d$Temperature)
pred.df$species <- ifelse(pred.df$pen,"S.pen","S.lyc")

pl <- ggplot(pred.df,aes(y=p.mean,ymin=p.PI.low,ymax=p.PI.high,x=species,fill=as.factor(temperature)))
pl <- pl + geom_bar(stat = "identity",position="dodge")
pl <- pl + geom_errorbar(width=0.5, position=position_dodge(width=0.9))
pl <- pl + scale_fill_discrete(name="temperature")
pl + ggtitle("predicted germination percentages")
```

### compare posterior estimates

Temperature difference in S.lyc
```{r}
S.lyc.temp <- p.link[,2]-p.link[,1]
dens(S.lyc.temp,show.HPDI = 0.95, show.zero = TRUE,xlim=c(-.1,.9),main="S.lyc 22 vs. 14")
```
Clearly temp influences S.lyc

Temperature difference in S.pen
```{r}
S.pen.temp <- p.link[,4]-p.link[,3]
dens(S.pen.temp,show.HPDI = 0.95, show.zero = TRUE, main="S.pen 22 vs. 14")
sum(S.pen.temp > 0) / length(S.pen.temp)
```
Less clear

S.lyc vs S.pen at 14
```{r}
S.pen.lyc.14 <- p.link[,3]-p.link[,1]
dens(S.pen.lyc.14,show.HPDI = 0.95, show.zero = TRUE, main="S.pen vs. S. lyc at 14˚")
sum(S.pen.lyc.14 > 0) / length(S.pen.lyc.14)
```
Clear difference

S.lyc vs S.pen at 22
```{r}
S.pen.lyc.22 <- p.link[,4]-p.link[,2]
dens(S.pen.lyc.22,show.HPDI = 0.95, show.zero = TRUE, main="S.pen vs. S. lyc at 22˚")
sum(S.pen.lyc.22 > 0) / length(S.pen.lyc.14)
```
No evidence for a difference