---
title: "Chapter 14 Problems"
author: "Julin N Maloof"
date: "3/8/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, autodep = TRUE)
```

## 14E1

Specify a distribition that the observations come from (actually each observation has a separate set of parameters).  Have the linear model use the distribution as predictors rather than the observations themselves

T_i ~ Poission(mu_i)
log(mu_i) = a + B*logPi_est
logPi_obs ~ dnorm(logPi_est_i, logPi_sigma_i)
A ~ dnorm(0,10)
B ~ dnorm(0,10)

## 14E2

## 14M1

They all come from the same distribution, with mean v and sd sigma_n

## 14M2

_Repeat WAIC comparison on milk primate data models from chapter 6, but include missing data (via imputation).  Does this affect model performance?_

### Without imputation

Updated code to use map2stan 

```{r}
# Rcode 6.21
library(rethinking)
data(milk)
d <- milk[ complete.cases(milk) , ]
d$neocortex <- d$neocortex.perc / 100
d$logmass <- log(d$mass)
dim(d)

d_list <- list(
kcal = d$kcal,
neocortex = d$neocortex,
logmass = d$logmass )

# Rcode 6.22

m6.11 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma) ,
    mu <- a,
    a ~ dnorm(0,100),
    sigma ~ dcauchy(0,1)),
  data=d_list , 
  chains = 4, cores = 1)

m6.12 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bn*neocortex,
    a ~ dnorm(0,100),
    bn ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
  data=d_list , 
  chains = 4, cores = 1)

m6.13 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bm*logmass,
    a ~ dnorm(0,100),
    bm ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
  data=d_list , 
  chains = 4, cores = 1)

m6.14 <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bn*neocortex + bm*logmass,
    a ~ dnorm(0,100),
    c(bn,bm) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
  data=d_list , 
  chains = 4, cores = 1)

#Rcode 6.24
( milk.models <- compare( m6.11 , m6.12 , m6.13 , m6.14 ) )

save.image("chapter14.Rdata")
```

### Now with missing data imputation

```{r}
library(rethinking)
load("chapter14.Rdata")
# Rcode 6.21
d.all <- milk
d.all$neocortex <- d.all$neocortex.perc / 100
d.all$logmass <- log(d.all$mass)
dim(d.all)

d.all_list <- list(
kcal = d.all$kcal,
neocortex = d.all$neocortex,
logmass = d.all$logmass )

# Rcode 6.22

m6.11.all <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma) ,
    mu <- a,
    a ~ dnorm(0,100),
    sigma ~ dcauchy(0,1)),
  data=d.all_list , 
  chains = 4, cores = 1)

m6.12.all <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bn*neocortex,
    a ~ dnorm(0,100),
    bn ~ dnorm(0,10),
    neocortex ~ dnorm(nu,sigma_nu),
    nu ~ dnorm(0.5,1),
    sigma_nu ~ dcauchy(0,1),
    sigma ~ dcauchy(0,1)),
  data=d.all_list , 
  chains = 4, cores = 1)

save.image("chapter14.Rdata")

m6.13.all <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bm*logmass,
    a ~ dnorm(0,100),
    bm ~ dnorm(0,10),
    neocortex ~ dnorm(nu,sigma_nu),
    nu ~ dnorm(0.5,1),
    sigma_nu ~ dcauchy(0,1),
    sigma ~ dcauchy(0,1)),
  data=d.all_list , 
  chains = 4, cores = 1)

save.image("chapter14.Rdata")

m6.14.all <- map2stan(
  alist(
    kcal ~ dnorm( mu , sigma ) ,
    mu <- a + bn*neocortex + bm*logmass,
    a ~ dnorm(0,100),
    c(bn,bm) ~ dnorm(0,10),
    neocortex ~ dnorm(nu,sigma_nu),
    nu ~ dnorm(0.5,1),
    sigma_nu ~ dcauchy(0,1),
    sigma ~ dcauchy(0,1)),
  data=d.all_list , 
  chains = 4, cores = 1)

save.image("chapter14.Rdata")
```

```{r}

precis(m6.11.all, depth=2)

precis(m6.12.all,depth=2)

precis(m6.13.all, depth=2)

precis(m6.14.all, depth=2)


#Rcode 6.24
( milk.models <- compare( m6.11 , m6.12 , m6.13 , m6.14 ) )
( milk.models.all <- compare( m6.11.all, m6.12.all , m6.13.all , m6.14.all ) )

```

This did not substantially change the ranking of the models.

## 14M3

_Repeat the divorce rate error models, but double the standard error.  Can you explain how doubling the errors affects the inference?_

original
```{r}
library(rethinking)
data("WaffleDivorce")
d <- WaffleDivorce
dlist <- list(
  div_obs=d$Divorce,
  div_sd=d$Divorce.SE,
  mar_obs=d$Marriage,
  mar_sd=d$Marriage.SE,
  A=d$MedianAgeMarriage )

m14.2 <- map2stan(
  alist(
    div_est ~ dnorm(mu,sigma),
    mu <- a + bA*A + bR*mar_est[i],
    div_obs ~ dnorm(div_est,div_sd),
    mar_obs ~ dnorm(mar_est,mar_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,10),
    bR ~ dnorm(0,10),
    sigma ~ dcauchy(0,2.5)
  ) ,
  data=dlist ,
  start=list(div_est=dlist$div_obs,mar_est=dlist$mar_obs) ,
  WAIC=FALSE , iter=5000 , warmup=1000, chains=3 , cores=3 ,
  control=list(adapt_delta=0.95) )
save.image("chapter14.Rdata")
```

double error
```{r}
library(rethinking)
load("chapter14.Rdata")
dlist <- list(
  div_obs=d$Divorce,
  div_sd=2*d$Divorce.SE,
  mar_obs=d$Marriage,
  mar_sd=d$Marriage.SE,
  A=d$MedianAgeMarriage )

m14.2.double <- map2stan(
  alist(
    div_est ~ dnorm(mu,sigma),
    mu <- a + bA*A + bR*mar_est[i],
    div_obs ~ dnorm(div_est,div_sd),
    mar_obs ~ dnorm(mar_est,mar_sd),
    a ~ dnorm(0,10),
    bA ~ dnorm(0,10),
    bR ~ dnorm(0,10),
    sigma ~ dcauchy(0,2.5)
  ) ,
  data=dlist ,
  start=list(div_est=dlist$div_obs,mar_est=dlist$mar_obs) ,
  WAIC=FALSE , iter=10000 , warmup=1000, chains=3 , cores=3 ,
  control=list(adapt_delta=0.95) )
```

```{r}
compare(m14.2,m14.2.double)
precis(m14.2,depth=2)
precis(m14.2.double,depth = 2)
```

## 14H1

_Fit a Poisson model predicting matings as a function of age in the elephants data set.  Age may not be measured acurately, so fit another model, including a standard error of +/-5 years.  Compare the models._

```{r 14H1_data}
library(rethinking)
data(elephants)
summary(elephants)
plot(elephants$AGE, elephants$MATINGS)
```
```{r 14H1_model1, results='hide'}
mean(elephants$MATINGS)
m14h1a <- map2stan(alist(
  MATINGS ~ dpois(lambda),
  log(lambda) <- a + b_age*AGE,
  a ~ dnorm(2.5,5),
  b_age ~ dnorm(0,1)
), data = elephants,
chains = 4,
cores = 2)
```

```{r 14h1a_plots}
plot(m14h1a)
precis(m14h1a)
exp(0.07)
exp(-1.54)

post <- link(m14h1a,data=data.frame(AGE=25:55))
head(post)
mu <- apply(post,2,mean)
PI <- apply(post,2,PI)

plot(x=25:55,y=mu,type="l")
points(elephants$AGE,elephants$MATINGS)
shade(PI,lim=25:55)
save.image("chapter14.Rdata")
```

model with error
```{r 14H1_model2, results='hide'}
elephants$AGE_SE <- 5
m14h1b <- map2stan(alist(
  MATINGS ~ dpois(lambda),
  log(lambda) <- a + b_age*AGE_est[i],
  a ~ dnorm(2.5,5),
  b_age ~ dnorm(0,1),
  AGE ~ dnorm(AGE_est,AGE_SE)
), data = elephants,
start=list(AGE_est=elephants$AGE),
chains = 4,
cores = 2)
```

```{r 14h1b_plots}
plot(m14h1b,ask=FALSE)
par(mfrow=c(1,1))
precis(m14h1b,depth = 2)
exp(0.07)
exp(-1.57)

post <- link(m14h1b,data=data.frame(AGE_est=25:55))
head(post)
mu.error <- apply(post,2,mean)
PI.error <- apply(post,2,PI)

plot(x=25:55,y=mu,type="l")
lines(x=25:55,y=mu.error,col="red")
points(elephants$AGE,elephants$MATINGS)
shade(PI.error,lim = 25:55)
shade(PI,lim=25:55)
```

Essentially no difference in the models

## 14H2

_expand error_

```{r 14H1_model3, results='hide'}
elephants$AGE_SE <- 20
m14h1c <- map2stan(alist(
  MATINGS ~ dpois(lambda),
  log(lambda) <- a + b_age*AGE_est[i],
  a ~ dnorm(2.5,5),
  b_age ~ dnorm(0,1),
  AGE ~ dnorm(AGE_est,AGE_SE)
), data = elephants,
start=list(AGE_est=elephants$AGE),
chains = 4,
cores = 2)
```

```{r 14h1c_plots}
plot(m14h1c,ask=FALSE)
par(mfrow=c(1,1))
precis(m14h1c)

post <- link(m14h1c,data=data.frame(AGE_est=25:55))
mu.error <- apply(post,2,mean)
PI.error <- apply(post,2,PI)

plot(x=25:55,y=mu,type="l")
lines(x=25:55,y=mu.error,col="red")
points(elephants$AGE,elephants$MATINGS)
shade(PI.error,lim = 25:55)
shade(PI,lim=25:55)
```

still a positive relationship with error of 20

```{r 14H1_model4, results='hide'}
elephants$AGE_SE <- 40
m14h1d <- map2stan(alist(
  MATINGS ~ dpois(lambda),
  log(lambda) <- a + b_age*AGE_est[i],
  a ~ dnorm(2.5,5),
  b_age ~ dnorm(0,1),
  AGE ~ dnorm(AGE_est,AGE_SE)
), data = elephants,
start=list(AGE_est=elephants$AGE),
chains = 4,
cores = 2)
```

```{r 14h1d_plots}
plot(m14h1d,ask=FALSE)
par(mfrow=c(1,1))
precis(m14h1d)

post <- link(m14h1d,data=data.frame(AGE_est=25:55))
mu.error <- apply(post,2,mean)
PI.error <- apply(post,2,PI)

plot(x=25:55,y=mu,type="l")
lines(x=25:55,y=mu.error,col="red")
points(elephants$AGE,elephants$MATINGS)
shade(PI.error,lim = 25:55)
shade(PI,lim=25:55)
```

```{r 14H1_model5, results='hide'}
elephants$AGE_SE <- 100
m14h1e <- map2stan(alist(
  MATINGS ~ dpois(lambda),
  log(lambda) <- a + b_age*AGE_est[i],
  a ~ dnorm(2.5,5),
  b_age ~ dnorm(0,1),
  AGE ~ dnorm(AGE_est,AGE_SE)
), data = elephants,
start=list(AGE_est=elephants$AGE),
chains = 4, 
iter = 5000,
warmup = 1000,
cores = 2)
```

```{r 14h1e_plots}
plot(m14h1e,ask=FALSE)
par(mfrow=c(1,1))
precis(m14h1e)

post <- link(m14h1e,data=data.frame(AGE_est=25:55))
mu.error <- apply(post,2,mean)
PI.error <- apply(post,2,PI)

plot(x=25:55,y=mu,type="l")
lines(x=25:55,y=mu.error,col="red")
points(elephants$AGE,elephants$MATINGS)
shade(PI.error,lim = 25:55)
shade(PI,lim=25:55)
```

at SE of 100 we no longer see an AGE effect, although note that the model is not being fit well.