---
title: "Chapter 12 Problems"
author: "Julin Maloof"
date: "10/27/2016"
output: 
  html_document: 
    keep_md: yes
---

## setup

```{r}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE)
library(rethinking)
library(brms)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## 12E1

The prior of normal(0,1) will provide more shrinkage

## 12E2

Instead of 

a_group ~ Normal(0,10)

use 

a_group ~ normal(a,sigma)
a ~ (0,10)
sigma ~ cacuhy(0,1)

## 12M1

alpha only

```{r}
data(reedfrogs, results='hide')
d <- reedfrogs
d$tank <- 1:nrow(d)
m12m1.tank <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1)
  ), data=d , iter=4000 , chains=4 )
```

```{r}
plot(m12m1.tank,ask=FALSE)
precis(m12m1.tank)
```


with predation

```{r, results='hide'}
d$pred2 <- ifelse(d$pred=="pred",1,0)
m12m1.tank.pred <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] + b_pred*pred2 ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1),
    b_pred ~ dnorm(0,5)
  ), data=d , iter=4000 , chains=4 )
```

```{r}
plot(m12m1.tank.pred,ask=FALSE)
precis(m12m1.tank.pred)
```

with size

```{r, results='hide'}
d$big <- ifelse(d$size=="big",1,0)
m12m1.tank.size <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] + b_big*big ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1),
    b_big ~ dnorm(0,5)
  ), data=d , iter=4000 , chains=4 )
```

```{r}
plot(m12m1.tank.size,ask=FALSE)
precis(m12m1.tank.size)
```

additive, with pred and size

```{r, results='hide'}
m12m1.tank.pred.size <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] + b_big*big + b_pred*pred2,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1),
    c(b_big,b_pred) ~ dnorm(0,5)
  ), data=d , iter=4000 , chains=4 )
```

```{r}
plot(m12m1.tank.pred.size,ask=FALSE)
precis(m12m1.tank.pred.size)
```

interaction, with pred and size

```{r, results='hide'}
m12m1.tank.pred.size.int <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] + b_big*big + b_pred*pred2 + b_big_pred*big*pred2,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1),
    c(b_big,b_pred,b_big_pred) ~ dnorm(0,5)
  ), data=d , iter=4000 , chains=4 )
```

```{r}
plot(m12m1.tank.pred.size.int,ask=FALSE)
precis(m12m1.tank.pred.size.int)
par(mfrow=c(1,1))
```

_Focus on the inferred variation across tanks.  Explain why it changes as it does across models_

At first pass we can just look at the `sigma` parameter from each model as this is the estimate of adaptive estimate of standard deviation from tank to tank.

```{r}
precis(m12m1.tank)
precis(m12m1.tank.pred)
precis(m12m1.tank.size)
precis(m12m1.tank.pred.size)
precis(m12m1.tank.pred.size.int)
```

Basically we see that having predation in the model reduces variance among tanks.  This is because predation is a strong predicor of survival, so including it in the model reduces the otherwise unexplained tank to tank variance.

## 12M2

_Compare the models you fit just above, using WAIC.  Can you reconcile the differences in WAIC with the posterior distributions of the models?_

```{r}
compare(m12m1.tank,m12m1.tank.pred,m12m1.tank.size,m12m1.tank.pred.size,m12m1.tank.pred.size.int)
```

Models that include `pred` have a smaller number of effective parameters and a lower WAIC.  This makes sense w.r.t. the posterior distributions; tanks 

## Fit one of these with brms

```{r, results='hide'}
m12m1.tank.pred.size.int.b <- 
  brm(surv | trials(density) ~ 0 + (1| tank) + pred*size,
               data=d,
               family=binomial(link = "logit"),
               prior=c(set_prior("cauchy(0,1)", class = "sd"),
                       set_prior("normal(0,5)", class = "b")))
```

```{r}
plot(m12m1.tank.pred.size.int.b)
m12m1.tank.pred.size.int.b
precis(m12m1.tank.pred.size.int)
```

