---
title: "Chapter-05-part2-assignment"
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name: Julin Maloof__


# For <del>04/18/2016</del> 4/22/16

## 5M2

My example of a masked relationship is predicting energy usage (kW pulled from PG&E) for the Brady/Koch househould.  Because they have solar, this should be predictable by number pf sunny hours in the day.  However, long, sunny days also correlate with heat, and they have an air conditioner, which will affect energy use in the opposite way.

## 5H1

```{r}
library(rethinking)
data(foxes)
data <- foxes
summary(data)
head(data)

m5h1.a <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.area*area,
    a <- dnorm(4.5, 2),
    b.area <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h1.a)
plot(precis(m5h1.a))

pred.data.5h1.a <- data.frame(area=seq(0,10,length.out=100))

mu.5h1.a <- link(m5h1.a,pred.data.5h1.a,1e4)
mu.5h1.a.mean <- apply(mu.5h1.a,2,mean)
mu.5h1.a.PI <- apply(mu.5h1.a,2,PI)

plot(weight~area,data=data)
lines(pred.data.5h1.a$area,mu.5h1.a.mean)
lines(pred.data.5h1.a$area,mu.5h1.a.PI[1,],lty=2)
lines(pred.data.5h1.a$area,mu.5h1.a.PI[2,],lty=2)

m5h1.b <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.groupsize*groupsize,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h1.b)
plot(precis(m5h1.b))

pred.data.5h1.b <- data.frame(groupsize=seq(0,10,length.out=100))

mu.5h1.b <- link(m5h1.b,pred.data.5h1.b,1e4)
mu.5h1.b.mean <- apply(mu.5h1.b,2,mean)
mu.5h1.b.PI <- apply(mu.5h1.b,2,PI)

plot(weight~groupsize,data=data)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.mean)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.PI[1,],lty=2)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.PI[2,],lty=2)
```

Neither variable is particularly important for predicting body weight

## 5H2
```{r}
m5h2 <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.area*area + b.groupsize*groupsize,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    b.area <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h2)
plot(precis(m5h2))

pred.data.5h2.gs <- data.frame(groupsize=seq(0,10,length.out=100),
                                      area=mean(data$area))

mu.5h2.gs <- link(m5h2,pred.data.5h2.gs,1e4)
mu.5h2.gs.mean <- apply(mu.5h2.gs,2,mean)
mu.5h2.gs.PI <- apply(mu.5h2.gs,2,PI)

plot(weight~groupsize,data=data,type="n")
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.mean)
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.PI[1,],lty=2)
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.PI[2,],lty=2)

pred.data.5h2.area <- data.frame(groupsize=mean(data$groupsize),
                                      area=seq(0,10,length.out=100))

mu.5h2.area <- link(m5h2,pred.data.5h2.area,1e4)
mu.5h2.area.mean <- apply(mu.5h2.area,2,mean)
mu.5h2.area.PI <- apply(mu.5h2.area,2,PI)

plot(weight~area,data=data,type="n")
lines(pred.data.5h2.area$area,mu.5h2.area.mean)
lines(pred.data.5h2.area$area,mu.5h2.area.PI[1,],lty=2)
lines(pred.data.5h2.area$area,mu.5h2.area.PI[2,],lty=2)

plot(groupsize~area,data=data)
```

## 5H3

#### Avg Food and Group Size

```{r 5H3_part1}
m5h3a <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.avgfood*avgfood + b.groupsize*groupsize,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    b.avgfood <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h3a)
plot(precis(m5h3a))

pred.data.5h3.gs <- data.frame(groupsize=seq(0,10,length.out=100),
                                      avgfood=mean(data$avgfood))

mu.5h3.gs <- link(m5h3a,pred.data.5h3.gs,1e4)
mu.5h3.gs.mean <- apply(mu.5h3.gs,2,mean)
mu.5h3.gs.PI <- apply(mu.5h3.gs,2,PI)

plot(weight~groupsize,data=data,type="n")
lines(pred.data.5h3.gs$groupsize,mu.5h3.gs.mean)
lines(pred.data.5h3.gs$groupsize,mu.5h3.gs.PI[1,],lty=2)
lines(pred.data.5h3.gs$groupsize,mu.5h3.gs.PI[2,],lty=2)

pred.data.5h3.avgfood <- data.frame(groupsize=mean(data$groupsize),
                                      avgfood=seq(0,2,length.out=100))

mu.5h3.avgfood <- link(m5h3a,pred.data.5h3.avgfood,1e4)
mu.5h3.avgfood.mean <- apply(mu.5h3.avgfood,2,mean)
mu.5h3.avgfood.PI <- apply(mu.5h3.avgfood,2,PI)

plot(weight~avgfood,data=data,type="n")
lines(pred.data.5h3.avgfood$avgfood,mu.5h3.avgfood.mean)
lines(pred.data.5h3.avgfood$avgfood,mu.5h3.avgfood.PI[1,],lty=2)
lines(pred.data.5h3.avgfood$avgfood,mu.5h3.avgfood.PI[2,],lty=2)

plot(groupsize~avgfood,data=data)
cor(data$groupsize,data$avgfood)
```
So Avgfood and GroupSize are correlated by have opposite effects on weight so they "mask" one another.  The model fits well when both are included.

#### Average Food, GroupSize, Area
```{r 5H3_part2}
m5h3b <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.avgfood*avgfood + b.groupsize*groupsize + b.area*area,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    b.avgfood <- dnorm(0,5),
    b.area <- dnorm(0.5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h3b)
plot(precis(m5h3b))
```

Now the coefficients for avgfood and area are both lower and could be 0

Question: is average food or area a better predictor of weight?

The three-predictor model isn't very helpful because the two predictors are corelated
```{r}
plot(area ~ avgfood, data=data)
cor(data$area, data$avgfood)
```

One possiblity is to look at the sigma in the two-way models.  Lowed sigma would indicate a better fit)
```{r}
#avgfood and groupsize
precis(m5h3a)

#area and groupsize
precis(m5h2)

```

They seem pretty similar to me; not sure how to ditinguish.

