---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name:__ _Julin Maloof_


# For 04/11/2016

## 5E1

_Which of the linear models below are multiple linear regressions?_

Two and Four

## 5E2

_Write down a multiple regression to evaluate the claim: Animal diversity is linearly related to latitude, but only after controlling for plant diversity_

I am not sure if we have full covered this, but:

$$
Diversity_i = \alpha + \beta_{lat}*lat_i + \beta_{plant}*plant_i
$$

## 5M1

_Invent your own example of a spurious correlation_

Air density is a predictor of Julin's average speed on the bicycle

Average miles ridden in the previous month is a predictor of Julin's speed on the bicycles

## 5M3

_How might a high divorce rate cause a high marriage rate? Can you think of a way to evaluate this using multiple regression?_

A high divorce rate could cause a high marriage rate because of there being more single people available for marriage

To test for this using multiple regression one could include a factor for the proprtion of marriage that are second marriages.

## 5M4

_Include %Mormons in the divorce rate model_

LDS data from the [wikipedia page](https://en.wikipedia.org/wiki/The_Church_of_Jesus_Christ_of_Latter-day_Saints_membership_statistics_(United_States)#Table)


First get the LDS data and combine it with the divorce data.  I cut and paste from chrome into excel and then did a little reformatting there before saving it as a .csv.

The `merge()` function is quite useful...
```{r 5M4}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce
LDS <- read.csv("LDS.csv")
head(LDS)
summary(LDS)
summary(d)
d <- merge(d,LDS,by.x="Location",by.y="State")
summary(d)
```

Take a look at the LDS data.
```{r}
hist(d$Percent.LDS)
hist(log10(d$Percent.LDS))
```

Very not normal...

Do some tranformation and standardiztion anyway
```{r}
d$Percent.LDS.s <- (d$Percent.LDS-mean(d$Percent.LDS)) / sd(d$Percent.LDS)
d$Percent.LDS.log10 <- log10(d$Percent.LDS)
d$Percent.LDS.log10.s <- (d$Percent.LDS.log10-mean(d$Percent.LDS.log10)) / sd(d$Percent.LDS.log10)
d$Marriage.s <- (d$Marriage - mean(d$Marriage))/sd(d$Marriage)
d$MedianAgeMarriage.s <- (d$MedianAgeMarriage - mean(d$MedianAgeMarriage)) / sd(d$MedianAgeMarriage)
```

Plots to see relationship between Divorce and LDS data
```{r}
plot(d$Divorce ~ d$Percent.LDS)
plot(d$Divorce ~ d$Percent.LDS.s)
plot(d$Divorce ~ d$Percent.LDS.log10)
plot(d$Divorce ~ d$Percent.LDS.log10.s)
```

None of this looks that great.

An alternative is to look at the residuals from the Marriage Rate + Age model and ask how those relate to the LDS data.

```{r}
m5M4.0 <- map(alist(
  Divorce ~ dnorm(mu, sigma),
  mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s,
  a ~ dnorm(10,10),
  bR ~ dnorm(0,1),
  bA ~ dnorm(0,1),
  sigma ~ dunif(0,20)),
  data=d)

precis(m5M4.0)

mu.m5M4.0 <- link(m5M4.0)
mu.m5M4.0.mean <- apply(mu.m5M4.0,2,mean)

mu.m5M4.0.resid <- d$Divorce - mu.m5M4.0.mean

plot(d$Percent.LDS,mu.m5M4.0.resid)
plot(d$Percent.LDS.log10,mu.m5M4.0.resid)
```


Neither are great; is there another transformation that would be better?  I can't really think of one.

```{r}
m5M4.1 <- map(alist(
  Divorce ~ dnorm(mu, sigma),
  mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s + bLDS*Percent.LDS.s,
  a ~ dnorm(10,10),
  bR ~ dnorm(0,1),
  bA ~ dnorm(0,1),
  bLDS ~ dnorm(0,1),
  sigma ~ dunif(0,20)),
  data=d)

precis(m5M4.1)

mu.m5M4.1 <- link(m5M4.1)
mu.m5M41.mean <- apply(mu.m5M4.1,2,mean)
mu.m5M41.PI <- apply(mu.m5M4.1,2,PI)
plot(d$Divorce, mu.m5M41.mean, xlab="Observed",ylab="predicted")
abline(a=0,b=1)
```

```{r}
m5M4.2 <- map(alist(
  Divorce ~ dnorm(mu, sigma),
  mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s + bLDS*Percent.LDS.log10.s,
  a ~ dnorm(10,10),
  bR ~ dnorm(0,1),
  bA ~ dnorm(0,1),
  bLDS ~ dnorm(0,1),
  sigma ~ dunif(0,20)),
  data=d)

precis(m5M4.2)

mu.m5M4.2 <- link(m5M4.2)
mu.m5M41.mean <- apply(mu.m5M4.2,2,mean)
mu.m5M41.PI <- apply(mu.m5M4.2,2,PI)
plot(d$Divorce, mu.m5M41.mean, xlab="Observed",ylab="predicted")
abline(a=0,b=1)
```

Overall the untransformed LDS data is a bit better and it does seem to have an impact.