---
title: "Chapter-05-part2-assignment"
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 4 problems

__Name: Julin Maloof__


# For 04/18/2016

## 5M2

My example of a masked relationship is predicting energy usage (kW pulled from PG&E) for the Brady/Koch househould.  Because they have solar, this should be predictable by number pf sunny hours in the day.  However, long, sunny days also correlate with heat, and they have an air conditioner, which will affect energy use in the opposite way.

## 5H1

```{r}
library(rethinking)
data(foxes)
data <- foxes
summary(data)
head(data)

m5h1.a <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.area*area,
    a <- dnorm(4.5, 2),
    b.area <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h1.a)
plot(precis(m5h1.a))

pred.data.5h1.a <- data.frame(area=seq(0,10,length.out=100))

mu.5h1.a <- link(m5h1.a,pred.data.5h1.a,1e4)
mu.5h1.a.mean <- apply(mu.5h1.a,2,mean)
mu.5h1.a.PI <- apply(mu.5h1.a,2,PI)

plot(weight~area,data=data)
lines(pred.data.5h1.a$area,mu.5h1.a.mean)
lines(pred.data.5h1.a$area,mu.5h1.a.PI[1,],lty=2)
lines(pred.data.5h1.a$area,mu.5h1.a.PI[2,],lty=2)

m5h1.b <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.groupsize*groupsize,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h1.b)
plot(precis(m5h1.b))

pred.data.5h1.b <- data.frame(groupsize=seq(0,10,length.out=100))

mu.5h1.b <- link(m5h1.b,pred.data.5h1.b,1e4)
mu.5h1.b.mean <- apply(mu.5h1.b,2,mean)
mu.5h1.b.PI <- apply(mu.5h1.b,2,PI)

plot(weight~groupsize,data=data)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.mean)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.PI[1,],lty=2)
lines(pred.data.5h1.b$groupsize,mu.5h1.b.PI[2,],lty=2)
```

Neither variable is particularly important for predicting body weight

## 5H2
```{r}
m5h2 <- map(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + b.area*area + b.groupsize*groupsize,
    a <- dnorm(4.5, 2),
    b.groupsize <- dnorm(0,5),
    b.area <- dnorm(0,5),
    sigma <- dunif(0,10)),
    data=data)

precis(m5h2)
plot(precis(m5h2))

pred.data.5h2.gs <- data.frame(groupsize=seq(0,10,length.out=100),
                                      area=mean(data$area))

mu.5h2.gs <- link(m5h2,pred.data.5h2.gs,1e4)
mu.5h2.gs.mean <- apply(mu.5h2.gs,2,mean)
mu.5h2.gs.PI <- apply(mu.5h2.gs,2,PI)

plot(weight~groupsize,data=data,type="n")
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.mean)
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.PI[1,],lty=2)
lines(pred.data.5h2.gs$groupsize,mu.5h2.gs.PI[2,],lty=2)

pred.data.5h2.area <- data.frame(groupsize=mean(data$groupsize),
                                      area=seq(0,10,length.out=100))

mu.5h2.area <- link(m5h2,pred.data.5h2.area,1e4)
mu.5h2.area.mean <- apply(mu.5h2.area,2,mean)
mu.5h2.area.PI <- apply(mu.5h2.area,2,PI)

plot(weight~area,data=data,type="n")
lines(pred.data.5h2.area$area,mu.5h2.area.mean)
lines(pred.data.5h2.area$area,mu.5h2.area.PI[1,],lty=2)
lines(pred.data.5h2.area$area,mu.5h2.area.PI[2,],lty=2)

plot(groupsize~area,data=data)
```

## 5H3
