---
title: "BRMS Intro and practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
```

## BRMS background

An alternative to map2stan for interfacting with [`Stan`](http://mc-stan.org/) is the [`brms`](https://github.com/paul-buerkner/brms) (Bayesian Regression Modelling using Stan) package.  While `map2stan` is a fantastic tool for learning model specification and being very explicit about the models being fit, `brms` has a more intuitive model specification and requires less data reformatting.  `brms` will be particularly familiar to those used to fitting models with `lmer`.

If you want to read someone else's introduction to `brms` you can do so [here](https://thinkinator.com/2016/01/12/r-users-will-now-inevitably-become-bayesians/).

You can also read the package author's README by scrolling down at the [github repo](https://github.com/paul-buerkner/brms)

Finally, the author provides a very detailed description available [here](https://cran.r-project.org/web/packages/brms/vignettes/brms_overview.pdf) or by typing `vignette("brms_overview")` after you have brms installed.

## Installation

Because `brms` is in rapid development I prefer to install the most recent version from github:

```{r install, eval=FALSE}
library(devtools)
install_github("paul-buerkner/brms")
```

(If you don't have `devtools` installed then you will first need to run `install.packages("devtools")`)

## Model specification in brms

Model specification is very similar to that in lmer, namely,

```{r model, eval=FALSE}
repsonse ~ predictor1 + (1|random_predictor2) ...
```

So to use the familiar `Tomato` dataset

```{r get_tomato}
library(brms)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
tomato <- read.csv("../Assignment_Chapter_09/TomatoR2CSHL.csv") #adjust path as necesary
summary(tomato)
head(tomato)
```

```{r fit_brms_example1, results='hide'}
hyp1 <- brm(hyp ~ trt + species + (1|shelf),
            data = tomato,
            prior = set_prior("normal(0,10)",class="b") # I'll explain priors in the next section
) 
```

Note:

* I did not have to specify priors for sigma or the random effects (although you can, see below)
* I did not have to create dummy variables for trt, species, or shelf !!
* The default number of chains is 4
```{r examine_brms_example1}
summary(hyp1)
plot(hyp1,ask=FALSE)
```

Note:

* The summary automatically displays informative coefficient names 

## Interactive model viewer

Run the following for cool interactive exploration of your posteriors
```{r, eval=FALSE}
launch_shiny(hyp1) #interactive graphical exploration
```


## Prior specification in brms

brms will automatically set priors for many model parameters, but it uses non-sensical priors (uniform across all real numbers) for main-effect coefficients, so __you must set priors for your main-effect coefficients__.  Most other priors have reasonable defaults.

The function for setting priors is `set_priors()`.  

Priors in brms have different classes.

* The "Intercept" class is for the overall intercept
* The "b" class is for coefficients on fixed effects (You can think of "b" as standing for "beta" coefficients)
* The "sigma" class is for the overall standard deviation.  Defaults to `student_t(3, 0, 10)`
* The "sd" class is for the standard deviation of random effects terms.  Defaults to `student_t(3, 0, 10)`

If you want to give the same prior to all members of a class, then just name the class.  In the example I gave above (repeated below) I am setting the priors for all beta coefficients in this case (trtL and the various species) to be normal(0,10).

```{r, eval=FALSE}
hyp1 <- brm(hyp ~ trt + species + (1|shelf),
            data = tomato,
            prior = set_prior("normal(0,10)",class="b") # I'll explain priors in the next section
) 
```

Multiple `set_prior` calls can be combined using `c()`.  If you want to specify different priors for specific coefficients, then you can name them in separate `set_prior` calls:

```{r, results='hide'}
hyp2 <- brm(hyp ~ trt + species + (1|shelf),
            data = tomato,
            prior = c(set_prior("normal(0,10)",class="b"), # sets prior for all b coefficients not otherwise specified
                      set_prior("normal(0,5)",class="b", coef = "trtL")) #sets a different prior for the "trtL" coefficient
) 
```

To see all priors that you could specify for a given model and how they are specified, use `get_prior()`

```{r}
get_prior(hyp ~ trt + species + (1|shelf),
          data = tomato)
```

Or, for a model you have already fit, you can see available priors and how they were set:

```{r}
get_prior(hyp2$formula,data=tomato)
```


Example of setting priors similar to Statistical Rethinking:

```{r, results='hide'}
hyp3 <- brm(hyp ~ trt + species + (1|shelf),
            data = tomato,
            prior = c(
              set_prior("normal(33,10",class="Intercept"), # prior for the intercept
              set_prior("normal(0,10)",class="b"), # sets prior for all b coefficients not otherwise specified
              set_prior("normal(0,5)",class="b", coef = "trtL"), #set prior for "trtL"
              set_prior("cauchy(0,1)", class = "sigma"), #half cauchy prior for sigma
              set_prior("normal(0,1)", class = "sd", group = "shelf") #prior for variation due to shelf
            ) 
)
```

More information can be found in section 4 of `vignette("brms_overview")` or by `?set_prior`

## Stan parameters

You can specify various Stan parameters similar to the way you would in map2stan:

```{r, eval=FALSE}
hyp4 <- brm(hyp ~ trt + species + (1|shelf),
            data = tomato,
            prior = set_prior("normal(0,10)",class="b"), 
            chains = 4, #the default anyway
            iter=4000,
            warmup = 1000,
            cores = 2
)
summary(hyp4)
plot(hyp4)
```

## Interactions

To specify interactions between predictors just replace the '+' with a '*':

```{r, results='hide'}
hyp5 <- brm(hyp ~ trt * species + (1|shelf), #trt by species interaction
            data = tomato,
            prior = set_prior("normal(0,10)",class="b"), 
)
```

```{r}
summary(hyp5)
plot(hyp5, ask=FALSE)
```


## Comparing models

Models can be compared using WAIC or another procedure, leave one out cross validation, using loo:

```{r}
waic(hyp1,hyp5)
loo(hyp1,hyp5)
```

So hyp5 is favored, but not by much.

At this point I am not sure how to do the weighting that can be done in map2stan

## Binomial and other models

To specify a non-Gaussian model use the `family` parameter in your call to brm.  I will illustrate with Liron's germination data.

```{r}
germination <- read.csv("../Assignment_Chapter_10/LironDataTime1.csv")
summary(germination)
head(germination)
germination$Germination2 <- ifelse(germination$Germination=="Yes",1,0) #need numeric response
germination$Temperature_b <- germination$Temperature - min(germination$Temperature) #baseline it
```

The `trials(1)` in the formula below indicates how many trials per row in the data frame.  Instead of a numeric this could also refer to a column in the data frame.

```{r}
germ1 <- brm(Germination2 | trials(1) ~ Species + Temperature_b,
             family = "binomial",
             prior=set_prior("normal(0,10)", class="b"),
             data = germination
             )
```

```{r}
summary(germ1)
plot(germ1, ask=FALSE)
```

Note: I did not have to specify the link function, `logit` is automatically used as a link function.

Available families (and there link functions) are described by typing `?family` and `?brmsfamily`

## Hypothesis testing

brms includes a function for Bayesian hypothesis testing.  For example to test if the effect of "TrtL" is non-zero (two-sided hypothesis test):

```{r}
summary(hyp1)
hypothesis(hyp1,"trtL = 0")
```

To test if the effect of "TrtL" is positive (one-sided hypothesis test):

```{r}
summary(hyp1)
hypothesis(hyp1,"trtL > 0")
```

To test if the coefficients for S.pennellii and S. habrochaites are different:

```{r}
hypothesis(hyp1, "speciesS.pennellii-speciesS.habrochaites = 0")
```

# Exercises

To gain familiarity with `brms` I am going to have you refit models from Statistical Rethinking with brms:

## Q1

## Q2

## Q3
