---
title: "BRMS homework"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, autodep = TRUE)
```

To gain familiarity with `brms` I am going to have you refit models from Statistical Rethinking with brms.  In all cases compare the brms and map2stan parameter estimates; they should be effectively the same if you have specified them correctly.

Setup:
```{r}
library(rethinking)
library(brms)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


## Q1

_Fit model 10.9 (R code 10.28) with brms.  Remember that you should not need to transform the predictors, create dummy variables, nor coerce indexes to use brms.  Compare to the map2stan fit.  Test whether the coefficient for gender is different from 0 in the brms model._

First using map2stan

```{r, results='hide'}
data(UCBadmit)
d <- UCBadmit
d$male <- ifelse( d$applicant.gender=="male" , 1 , 0 )
d$dept_id <- coerce_index( d$dept )
m10.9 <- map2stan(
  alist(
    admit ~ dbinom( applications , p ) ,
    logit(p) <- a[dept_id] + bm*male ,
    a[dept_id] ~ dnorm(0,10) ,
    bm ~ dnorm(0,10)
  ) , data=d, 
  chains=4,warmup = 1000,iter = 4000 )
```


now brms
```{r, results='hide'}
head(d)
m10.9.brms <- brm(admit | trials(applications) ~ 0 + dept + applicant.gender,
                  family = "binomial",
                  prior = set_prior("normal(0,10)", class="b"),
                  data=d,
                  warmup = 1000,
                  iter = 4000
)
```

```{r}
precis(m10.9,depth=2)
summary(m10.9.brms)
```

The fits are essentially identical.

To test the effect of gender:

```{r}
(hyp1 <- hypothesis(m10.9.brms, "applicant.gendermale = 0"))
plot(hyp1)
```

No evidence of a gender effect

## Q2

_Refit models 12.1 and 12.2 (Rcode 12.2 and 12.3) with brms_

Get the data

```{r}
data(reedfrogs)
d <- reedfrogs
# make the tank cluster variable
d$tank <- 1:nrow(d)
head(d)
```


m12.1 with map2stan
```{r, results='hide'}
m12.1 <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( 0 , 5 )
  ), data=d )
```

model m12.1 with brms

```{r, results='hide'}
m12.1.brms <- brm(surv | trials(density) ~ 0 + as.factor(tank),
                  prior = set_prior("normal(0,5)", class = "b"),
                  family = "binomial",
                  data=d)
```

```{r}
precis(m12.1, depth=2)
summary(m12.1.brms)
plot(coef(m12.1),fixef(m12.1.brms), xlab="map2stan", ylab="brms",main= "compare coefficients")
```

same estimates

```{r, results='hide'}
m12.2 <- map2stan(
  alist(
    surv ~ dbinom( density , p ) ,
    logit(p) <- a_tank[tank] ,
    a_tank[tank] ~ dnorm( a , sigma ) ,
    a ~ dnorm(0,1) ,
    sigma ~ dcauchy(0,1)
  ), data=d , iter=4000 , chains=4 )
```


In map2stan you can set a prior on the mean of the random effect.  I am not sure how to do that in brms, so I just include a normal intercept.
```{r, results='hide'}
m12.2.brms <- brm( surv | trials(density) ~ (1|tank),
                   family="binomial",
                   prior=c(
                     set_prior("normal(0,1)", class="Intercept"),
                     set_prior("cauchy(0,1)", class="sd")),
                   data=d,
                   iter=4000)
```

```{r}
print("map2stan")
precis(m12.2,depth=2)
print("----------")
print("brms")
summary(m12.2.brms)
```

To get the coefficients to line up I need to add the fixed effect intercept from brms to the ranef estimates
```{r}
plot(coef(m12.2)[1:48],ranef(m12.2.brms)$tank+fixef(m12.2.brms)[,"mean"],xlab="map2stan",ylab="brms")
```


## Q3

Use both brms and map2stan to fit a model for `intleng` as a function of `species`, `trt` and their interaction, and include `shelf` as a random effect.

Get the data and set it up for map2stan
```{r}
tomato <- read.csv("../Assignment_Chapter_09/TomatoR2CSHL.csv")
d.tomato <- tomato #for map2stan
d.tomato$species_id <- coerce_index(tomato$species)
d.tomato$shelf_id <- coerce_index(tomato$shelf)
d.tomato$shade <- ifelse(tomato$trt=="L",1,0)
d.tomato <- na.omit(d.tomato[,c("intleng","species_id","shelf_id","shade","species")])
mean(d.tomato$intleng)
```

map2stan fit
```{r, results='hide'}
mq3 <- map2stan(alist(
  intleng ~ dnorm(mu,sigma),
  mu <- a_species[species_id] + b_shade_species[species_id]*shade + a_shelf[shelf_id],
  sigma ~ dcauchy(0,1),
  a_species[species_id] ~ dnorm(20,10),
  a_shelf[shelf_id] ~ dnorm(0,sigma_shelf),
  sigma_shelf ~ dcauchy(0,1),
  b_shade_species[species_id] ~ dnorm(0,10)
),
data=d.tomato,
chains = 4,
iter=5000
)
```

brms fit

the prior specification gets ugly here becuase I am trying to match the individual intercept paramterization that I used for map2stan
```{r, results='hide'}
mq3.brms <- brm(intleng ~ 0 + species*trt + (1|shelf),
                prior = c(
                  set_prior("normal(0,10)",class = "b"), #for trtL and interactions
                  set_prior("normal(20,10)", class = "b", coef = "speciesS.chilense"),
                  set_prior("normal(20,10)", class = "b", coef = "speciesS.chmielewskii"),
                  set_prior("normal(20,10)", class = "b", coef = "speciesS.habrochaites"),
                  set_prior("normal(20,10)", class = "b", coef = "speciesS.pennellii"),
                  set_prior("normal(20,10)", class = "b", coef = "speciesS.peruvianum"),
                  
                  set_prior("cauchy(0,1)", class = "sigma"), #for overall sigma
                  set_prior("cauchy(0,1)", class = "sd") # for shelf
                ),
                data=tomato,
                iter=5000
)
```

```{r}
precis(mq3,depth=2)
summary(mq3.brms)
ranef(mq3.brms)
```

The shelf effects are too large in the map2stan model.

## Try with intercept being equal to S.chilense

```{r}
library(reshape2)
d.tomato$index <- 1:nrow(d.tomato)
d.tomato2 <- dcast(d.tomato,index + intleng + shelf_id + shade ~ species, value.var="species",fun.aggregate = length )
colnames(d.tomato2) <- sub(". ","_",colnames(d.tomato2),fixed=T)
head(d.tomato2)
```
map2stan fit
```{r, results='hide'}
mq3.2 <- map2stan(alist(
  intleng ~ dnorm(mu,sigma),
  mu <- a + 
    b_chm * S_chmielewskii +
    b_hab * S_habrochaites +
    b_pen * S_pennellii +
    b_per * S_peruvianum +
    b_shade * shade + 
    b_chm_shade * S_chmielewskii * shade +
    b_hab_shade * S_habrochaites * shade +
    b_pen_shade * S_pennellii * shade +
    b_per_shade * S_peruvianum * shade +    
    a_shelf[shelf_id],
  a ~ dnorm(20,10),
  c(b_chm,b_hab,b_pen,b_per,b_shade,b_chm_shade,b_hab_shade,b_pen_shade,b_per_shade) ~ dnorm(0,10),
  sigma ~ dcauchy(0,1),
  a_shelf[shelf_id] ~ dnorm(0,sigma_shelf),
  sigma_shelf ~ dcauchy(0,1)
),
data=d.tomato2,
chains = 4,
iter=5000
)
```

brms fit

```{r, results='hide'}
mq3.2.brms <- brm(intleng ~ + species*trt + (1|shelf),
                prior = c(
                  set_prior("normal(20,10)",class = "Intercept"),
                  set_prior("normal(0,10)",class = "b"), #for species, shade, and interactions
                  set_prior("cauchy(0,1)", class = "sigma"), #for overall sigma
                  set_prior("cauchy(0,1)", class = "sd") # for shelf
                ),
                data=tomato,
                iter=5000
)
```

```{r}
print("map2stan")
precis(mq3.2,depth=2)
print("----------------")
print("brms")
summary(mq3.2.brms)
ranef(mq3.2.brms)
```

Okay, these are really similar.  Not sure what is going wrong with the earlier specification.