---
title: "Tomato Problem"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,autodep = TRUE)
```

## Problem from Julin

Do a Bayesian analysis of hypocotyl length (hyp) in the attached data sheet.


1) Does the best model include species, trt, or both?

2) Evaluate the hypothesis that trt has an effect on hypocotyl length

If you get stuck early, don't worry, come anyway abd we can work on it in R Club

## Load libraries and get the data
```{r}
library(rethinking)
library(ggplot2)
library(reshape2)
data <- read.csv("TomatoR2CSHL.csv")
head(data)
summary(data)
```

## take a look at the data: boxplot
```{r}
pl <- ggplot(data,aes(y=hyp,x=species,fill=trt))
pl <- pl + geom_boxplot(position="dodge")
pl 
```

## take a look at the data: histograms
```{r}
pl <- ggplot(data,aes(x=hyp,fill=trt))
pl <- pl + geom_density()
pl <- pl + facet_grid(species ~ trt)
pl
```

## fit a model with trt
```{r, results='hide'}
mean(data$hyp) #33.35
data.trt <- data[,c("hyp","trt")]
data.trt$trt <- as.numeric(data$trt)-1
m.trt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt,
    a ~ dnorm(mu=33,sd=10),
    bT ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.trt,chains = 4,cores = 1)
```

```{r}
plot(m.trt)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.trt)
precis(m.trt)
plot(precis(m.trt))
```

## fit a model with species
```{r, results='hide'}
data.species <- data[,c("hyp","species")]
data.species$id <- 1:nrow(data.species)
data.species <- dcast(data.species, hyp + id ~ species, value.var="species", fun.aggregate = length)
colnames(data.species) <- sub(". ","_",fixed = TRUE, colnames(data.species))
head(data.species)
data.species <- data.species[,c(-2,-4)]
m.species <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species, chains=4, cores = 1)
```

```{r}
plot(m.species)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species)
precis(m.species)
plot(precis(m.species))
```

## fit a model with trt and species
```{r, results='hide'}
data.species.trt <- data[,c("hyp","species","trt")]
data.species.trt$id <- 1:nrow(data.species.trt)
data.species.trt <- dcast(data.species.trt, hyp + trt + id ~ species, value.var="species", fun.aggregate = length)
colnames(data.species.trt) <- sub(". ","_",fixed = TRUE, colnames(data.species.trt))
head(data.species.trt)
data.species.trt <- data.species.trt[,c(-3,-5)]
data.species.trt$trt <- as.numeric(data.species.trt$trt)-1
head(data.species.trt)
m.species.trt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt, chains=4, cores = 1)

```

```{r}
plot(m.species.trt)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt)
precis(m.species.trt)
plot(precis(m.species.trt))
```

## compare models
```{r}
compare(m.trt,m.species,m.species.trt)
plot(compare(m.trt,m.species,m.species.trt))
coeftab(m.trt,m.species,m.species.trt)
plot(coeftab(m.trt,m.species,m.species.trt))
```

There is strong support for the model that includes both species and trt

This in and of itself is an indication that trt effects hypocotyl length.  Looking more closely, the 95% confidence intervals for bT are way above 0:
```{r}
precis(m.species.trt,prob = .95)
```

And none of the posterior samples for bT are less than or equal to zero.
```{r}
bT.post <- extract.samples(m.species.trt)$bT
dens(bT.post, show.zero = TRUE, show.HPDI = 0.95)
sum(bT.post<=0) / length(bT.post) * 100
```

# Try sqrt transformation


## fit a model with trt and species, with sqrt tranformation
```{r, results='hide'}
data.species.trt.sqrt <- within(data.species.trt, hyp <- sqrt(hyp))

head(data.species.trt.sqrt)
m.species.trt.sqrt <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=5.7,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt.sqrt, chains=4, cores = 1)
```

```{r}
plot(m.species.trt.sqrt)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt.sqrt)
precis(m.species.trt.sqrt)
round(precis(m.species.trt.sqrt)@output[,1:4]^2,3)
precis(m.species.trt)
plot(precis(m.species.trt.sqrt))
```

## compare models
```{r}
compare(m.species.trt,m.species.trt.sqrt)
```

There is strong support for the model that includes both species and trt

This in and of itself is an indication that trt effects hypocotyl length.  Looking more closely, the 95% confidence intervals for bT are way above 0:
```{r}
precis(m.species.trt.sqrt,prob = .95)
```

And none of the posterior samples for bT are less than or equal to zero.
```{r}
bT.post <- extract.samples(m.species.trt.sqrt)$bT
dens(bT.post, show.zero = TRUE, show.HPDI = 0.95)
sum(bT.post<=0) / length(bT.post) * 100
```

## Is it really fair to compare the transformed and non-transformed models by WAIC?  Maybe the scale in and of itself matters?  Try fitting when data is divided by 10

```{r, results='hide'}
data.species.trt.10 <- within(data.species.trt, hyp <- hyp/10)

head(data.species.trt.10)
m.species.trt.10 <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + bT * trt + bChi * S_chilense + bHab * S_habrochaites + bPen * S_pennellii + bPer * S_peruvianum,
    a ~ dnorm(mu=3.3,sd=10),
    c(bChi, bHab, bPen, bPer, bT) ~ dnorm(0,1),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt.10, chains=4, cores = 1)
```

```{r}
plot(m.species.trt.10)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt.10)
precis(m.species.trt.10)
plot(precis(m.species.trt.10))
```

## compare models
```{r}
compare(m.species.trt,m.species.trt.10,m.species.trt.sqrt)
```

# model with interaction

```{r, results='hide'}
m.species.trt.int <- map2stan(
  alist(
    hyp ~ dnorm(mu,sigma),
    mu <- a + 
      bT * trt + 
      bChi * S_chilense + 
      bHab * S_habrochaites + 
      bPen * S_pennellii + 
      bPer * S_peruvianum +
      bChi_T * S_chilense * trt + 
      bHab_T * S_habrochaites * trt + 
      bPen_T * S_pennellii * trt + 
      bPer_T * S_peruvianum * trt
      ,
    a ~ dnorm(mu=33,sd=10),
    c(bChi, bHab, bPen, bPer, bT, bChi_T, bHab_T, bPen_T, bPer_T) ~ dnorm(0,10),
    sigma ~ dcauchy(0,1)),
    data=data.species.trt, chains=4, cores = 1)
```

```{r}
plot(m.species.trt.int)
par(mfrow=c(1,1),mfcol=c(1,1))
pairs(m.species.trt.int)
precis(m.species.trt.int)
plot(precis(m.species.trt.int))
```

## compare models
```{r}
compare(m.trt,m.species,m.species.trt,m.species.trt.int)
plot(compare(m.trt,m.species,m.species.trt,m.species.trt.int))
coeftab(m.trt,m.species,m.species.trt,m.species.trt.int)
plot(coeftab(m.trt,m.species,m.species.trt,m.species.trt.int))
```

## probability masses



## predictions...