# Statistical Rethinking Chapter 3 problems

__Name:__ _Julin Maloof_

## Set up

```{r set up}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
library(rethinking)
```

## 3E1
How much posterior probability lies below _p_ = 0.2?  
For this we count the number of samples with _p_ < 0.2 and compare that to the total number of samples
```{r }
sum(samples < 0.2) / length(samples)
```

## 3E2
How much posterior probability lies above _p_ = 0.8?
```{r}
sum(samples > 0.8) / length(samples)
```

## 3E3
How much posterior probability lies between 0.2 and 0.8?
```{r}
sum(samples > 0.2 & samples < 0.8) / length(samples)
```

## 3E4
20% of the posterior probability lies below which value of p?
```{r}
quantile(samples,.2)
```

## 3E5
20% of the posterior probability lies above which value of p?
```{r}
quantile(samples,.8)
```

## 3E6
Which values of _p_ contain the narrowest interval equal to 66% of the posterior probability?
```{r}
HPDI(samples,0.66)
```

## 3E7
Which calues of p contain 66% of the posterior probability, assuming equal posterior probabilities both below and above the interval?
```{r}
PI(samples,0.66)
```

## 3M1

```{r}
p_grid <- seq(0,1,length.out=10000)
prior <- rep(1,length(p_grid))
unstd.posterior <- dbinom(8,15,p_grid) * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior,type="l")
```

## 3M2

```{r}
samples <- sample(p_grid,size=10000,prob=posterior,replace=TRUE)
library(ggplot2)
qplot(samples,geom="density") + xlab("p")
HPDI(samples,p=0.9)
```


_STOP AFTER 3M2 FOR 02/25 ASSIGNMENT_

## 3M3

## 3M4

## 3M5

## 3H1

## 3H2

## 3H3

## 3H4

## 3H5