---
output: 
  html_document: 
    keep_md: yes
---
# Statistical Rethinking Chapter 3 problems

__Name:__ _Julin Maloof_

## Set up

```{r set up}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
library(rethinking)
```

## 3E1
How much posterior probability lies below _p_ = 0.2?  
For this we count the number of samples with _p_ < 0.2 and compare that to the total number of samples
```{r }
sum(samples < 0.2) / length(samples)
```

## 3E2
How much posterior probability lies above _p_ = 0.8?
```{r}
sum(samples > 0.8) / length(samples)
```

## 3E3
How much posterior probability lies between 0.2 and 0.8?
```{r}
sum(samples > 0.2 & samples < 0.8) / length(samples)
```

## 3E4
20% of the posterior probability lies below which value of p?
```{r}
quantile(samples,.2)
```

## 3E5
20% of the posterior probability lies above which value of p?
```{r}
quantile(samples,.8)
```

## 3E6
Which values of _p_ contain the narrowest interval equal to 66% of the posterior probability?
```{r}
HPDI(samples,0.66)
```

## 3E7
Which values of p contain 66% of the posterior probability, assuming equal posterior probabilities both below and above the interval?
```{r}
PI(samples,0.66)
```

## 3M1

```{r}
p_grid <- seq(0,1,length.out=10000)
prior <- rep(1,length(p_grid))
unstd.posterior <- dbinom(8,15,p_grid) * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior,type="l")
```

## 3M2

```{r}
samples <- sample(p_grid,size=10000,prob=posterior,replace=TRUE)
library(ggplot2)
qplot(samples,geom="density") + xlab("p")
HPDI(samples,p=0.9)
```

## 3M3

Construct a posterior predictive check for the model and data in 3M1 and 3M2.

We have sampled from the posterior distribution.  We can use these samples to weight the probabilities in rbinom
```{r}
water_simulation <- rbinom(1e5,size=15,prob=samples)
head(water_simulation) #number of water in 15 tosses
```

To calculate the probability of 8 waters in 15 tosses we can just ask what proportion of the samples have a value of 8.

```{r}
sum(water_simulation==8) / length(water_simulation)
simplehist(water_simulation)
```

## 3M4

Using the new posterior distribution (3M1) calculate the probability of 6 waters in 9 tosses.

```{r}
water_simulation2 <- rbinom(1e5, size = 9, prob = samples)
sum(water_simulation2==6) / length(water_simulation2)
simplehist(water_simulation2)
```

## 3M5

now use a prior with 0 below p = 0.5 and constant above

```{r}
p_grid <- seq(0,1,length.out=10000)
prior <- ifelse(p_grid < 0.5 , 0, 1)
unstd.posterior <- dbinom(8,15,p_grid) * prior
posterior2 <- unstd.posterior / sum(unstd.posterior)
plot(p_grid,posterior2,type="l")
samples2 <- sample(p_grid,size=10000,prob=posterior2,replace=TRUE)
library(ggplot2)
qplot(samples2,geom="density") + xlab("p")
HPDI(samples2,p=0.9)
```

```{r}
water_simulation <- rbinom(1e5,size=15,prob=samples2)
sum(water_simulation==8) / length(water_simulation)
simplehist(water_simulation)
```

```{r}
water_simulation2 <- rbinom(1e5, size = 9, prob = samples2)
sum(water_simulation2==6) / length(water_simulation2)
simplehist(water_simulation2)
```

But should I be looking directly at the posterior?
```{r}
median(samples)
chainmode(samples,adj=0.1)
p_grid[which.max(posterior)]
median(samples2)
chainmode(samples2,adj=0.1)
p_grid[which.max(posterior2)]
```

# H problems

Set up
```{r}
library(rethinking)
data(homeworkch3)
```

## 3H1
use grid approximation to compute the posterior distribution for the probability of a birth being a boy
```{r}
p_grid <- seq(0,1,length.out = 10000)
prior <- rep(1,length(p_grid))
likelihood <- dbinom(sum(birth1,birth2),
                     size = length(c(birth1,birth2)),
                     prob = p_grid)
unstd.posterior <- prior * likelihood
posterior <- unstd.posterior/sum(unstd.posterior)
plot(p_grid,posterior,type="l")
print(paste("The posterior probability is maximized at p = ",round(p_grid[which.max(posterior)],3)))
```

## 3H2

```{r}
samples <- sample(p_grid,size = 1e4,prob=posterior,replace=TRUE)
HPDI(samples,c(.5,.89,.97))
```

## 3H3

```{r}
boys.predicted <- rbinom(1e4,size = 200,prob=samples)
head(boys.predicted)
dens(boys.predicted,show.HPDI = .95)
abline(v=111)
```
The model fits the data well

## 3H4

I am not entirely sure what is being asked here.  I think the idea is that we the current model to predict first borns.  The assumption is that the proportion of boys should be the same in first borns and second borns.

```{r}
firstborn.boys.predicted <- rbinom(1e4,size = 100,prob=samples)
dens(firstborn.boys.predicted,show.HPDI = .95,adj=1)
# how many boys are in first born?
sum(birth1)
abline(v=51)
```
Not so good!  There are many fewer first born boys than the model predicts

## 3H5

Number of firstborn girls
```{r}
first.girls <- length(birth1) - sum(birth1)
first.girls
```

number of firstborn girls followed by boys
```{r}
boys.follow.girls <- sum(birth1==0 & birth2==1)
boys.follow.girls
```

predicted boys in a sample of 49

```{r}
boys.follow.girls.prediction <- rbinom(1e4,size=49,prob=samples)
median(boys.follow.girls.prediction)
dens(boys.follow.girls.prediction,show.HPDI = .95,adj=1)
abline(v=boys.follow.girls)
```
The model does very poorly.  Families that have girls first are more likely to have boys second.  Why?  See [this paper](http://www.bmj.com/content/338/bmj.b1211)
